---
title: "Hybridization_AmphibianWebScrape"
author: "Jadyn Sethna"
date: "2025-02-13"
output: html_document
---

#Set-Up
Download necessary packages for webscraping here. You will need to use the function install.packages() if you haven't used these packages before. 

You can learn more about rvest here: https://rvest.tidyverse.org


```{r setup, include=FALSE}
library(rvest) #package used for webscraping html text. 
library(httr)
library(xml2)
library(tidyverse) #package used for datacleaning,contains lots of common packages
library(polite)
library(dplyr) #package used for datacleaning and organization
library(stringr) #package to remove excess space and adjust formatting when pulling data from a web page

```

#Webscraping

You can Access the Amphibians of the World Dataframe here: 
https://amphibiansoftheworld.amnh.org

```{r webscrape}

library(rvest)
url = 'https://amphibiansoftheworld.amnh.org/content/search?subtree=&subtree_id=&country%5B%5D=563&geo-resident=true&geo-likely-occurs=true&geo-endemic=true&geo-introduced=true&search_type=count'

web_amphib<-read_html(url) #Rvest package

# Extract all the text from the page
amphib_text <- web_amphib%>% html_nodes("body") %>% html_text()

# Extract all links on the page
links <- web_amphib %>% html_nodes("a") %>% html_attr("href")

amphib<-web_amphib %>%html_elements("head")

#####Okay it seems to be working I'm getting amphibian data####### Now start filtering

#Use the inspect devtools to see the structure of the amphibians of the wold database. Countries are listed under option select this node for a list of options. 
# Extract all country names from the <option> tags

countries <- web_amphib %>%
  html_nodes("option") %>%  # Select all <option> elements
  html_text()  # Extract the text (country name
print(countries)


```

###Canada Dataframe
IT WORKS! DO NOT TOUCH THIS BLOCK
```{r loop data pull}

#example URL for Canada

base_url='https://amphibiansoftheworld.amnh.org/content/search?subtree=&subtree_id=&country%5B%5D=563&geo-resident=true&geo-likely-occurs=true&geo-endemic=true&geo-introduced=true&search_type=count'

page_content=read_html(base_url)


# Find all species blocks
species_blocks <- page_content %>%
  html_nodes("div.Species")  # Adjust this selector if necessary

# Initialize a list to store the data for each species
species_data <- list()

# Loop through each species block and extract both species name and location data
for (i in seq_along(species_blocks)) {
  # Extract the species name (assumed to be inside an <a> tag within the species block)
  species_name <- species_blocks[i] %>%
    html_node("a") %>%
    html_text() %>%
    trimws()
  
  # Extract location data (assumed to be inside <div> with class "highlight")
  location_nodes <- species_blocks[i] %>%
    html_nodes("div.highlight") %>%
    html_text() %>%
    trimws()

  # Initialize variables to store locations by type
  natural_resident <- character(0)
  introduced <- character(0)
  endemic <- character(0)
  likely <- character(0)
  
  # Loop through each location node to check if it matches the types
  for (location in location_nodes) {
    location_cleaned <- str_replace_all(location, "\n", " ") %>%
      str_replace_all("\\s+", " ") %>%
      str_trim()

    # Check for keywords and assign locations accordingly
    if (grepl("Natural Resident", location_cleaned)) {
      natural_resident <- c(natural_resident, str_extract(location_cleaned, "(?<=Natural Resident: ).*"))
    }
    
    if (grepl("Introduced", location_cleaned)) {
      introduced <- c(introduced, str_extract(location_cleaned, "(?<=Introduced: ).*"))
    }
    
    if (grepl("Endemic", location_cleaned)) {
      endemic <- c(endemic, str_extract(location_cleaned, "(?<=Endemic: ).*"))
    }
    
    if (grepl("Likely", location_cleaned)) {
      likely <- c(likely, str_extract(location_cleaned, "(?<=Likely: ).*"))
    }
  }

  # Combine the extracted data into a data frame for each species
  species_data[[i]] <- data.frame(
    Species = species_name,
    Natural_Resident = paste(natural_resident, collapse = "; "),  # Join all locations if there are multiple
    Introduced = paste(introduced, collapse = "; "),
    Endemic = paste(endemic, collapse = "; "),
    Likely = paste(likely, collapse = "; "),
    stringsAsFactors = FALSE
  )
}

# Combine all species data into a single data frame
species_df <- bind_rows(species_data)

# Clean up any residual issues with whitespace
species_df <- species_df %>%
  mutate(
    Species = str_trim(Species),
    Natural_Resident = str_trim(Natural_Resident),
    Introduced = str_trim(Introduced),
    Endemic = str_trim(Endemic),
    Likely = str_trim(Likely)
  )

# View the final data frame
print(species_df)

# View the cleaned data frame
View(species_df)

CanadaAmphib<-species_df

write.csv(CanadaAmphib,'CanadaAmphibOnly.csv')


```



###Mexico Dataframe
IT WORKS! DO NOT TOUCH THIS BLOCK
```{r loop data pull}

#example URL for Mexico

base_url='https://amphibiansoftheworld.amnh.org/content/search?subtree=&subtree_id=&country%5B%5D=525&geo-resident=true&geo-likely-occurs=true&geo-endemic=true&geo-introduced=true&search_type=count'

page_content=read_html(base_url)


# Find all species blocks
species_blocks <- page_content %>%
  html_nodes("div.Species")  # Adjust this selector if necessary

# Initialize a list to store the data for each species
species_data <- list()

# Loop through each species block and extract both species name and location data
for (i in seq_along(species_blocks)) {
  # Extract the species name (assumed to be inside an <a> tag within the species block)
  species_name <- species_blocks[i] %>%
    html_node("a") %>%
    html_text() %>%
    trimws()
  
  # Extract location data (assumed to be inside <div> with class "highlight")
  location_nodes <- species_blocks[i] %>%
    html_nodes("div.highlight") %>%
    html_text() %>%
    trimws()

  # Initialize variables to store locations by type
  natural_resident <- character(0)
  introduced <- character(0)
  endemic <- character(0)
  likely <- character(0)
  
  # Loop through each location node to check if it matches the types
  for (location in location_nodes) {
    location_cleaned <- str_replace_all(location, "\n", " ") %>%
      str_replace_all("\\s+", " ") %>%
      str_trim()

    # Check for keywords and assign locations accordingly
    if (grepl("Natural Resident", location_cleaned)) {
      natural_resident <- c(natural_resident, str_extract(location_cleaned, "(?<=Natural Resident: ).*"))
    }
    
    if (grepl("Introduced", location_cleaned)) {
      introduced <- c(introduced, str_extract(location_cleaned, "(?<=Introduced: ).*"))
    }
    
    if (grepl("Endemic", location_cleaned)) {
      endemic <- c(endemic, str_extract(location_cleaned, "(?<=Endemic: ).*"))
    }
    
    if (grepl("Likely", location_cleaned)) {
      likely <- c(likely, str_extract(location_cleaned, "(?<=Likely: ).*"))
    }
  }

  # Combine the extracted data into a data frame for each species
  species_data[[i]] <- data.frame(
    Species = species_name,
    Natural_Resident = paste(natural_resident, collapse = "; "),  # Join all locations if there are multiple
    Introduced = paste(introduced, collapse = "; "),
    Endemic = paste(endemic, collapse = "; "),
    Likely = paste(likely, collapse = "; "),
    stringsAsFactors = FALSE
  )
}

# Combine all species data into a single data frame
species_df <- bind_rows(species_data)


# Clean up any residual issues with whitespace
species_df <- species_df %>%
  mutate(
    Species = str_trim(Species),
    Natural_Resident = str_trim(Natural_Resident),
    Introduced = str_trim(Introduced),
    Endemic = str_trim(Endemic),
    Likely = str_trim(Likely)
  )

# View the final data frame
print(species_df)

# View the cleaned data frame
View(species_df)

MexicoAmphib<-species_df
write.csv(MexicoAmphib,'MexicoAmphibOnly.csv')
```




###USA Dataframe
IT WORKS! DO NOT TOUCH THIS BLOCK
```{r loop data pull}

#example URL for USA

base_url='https://amphibiansoftheworld.amnh.org/content/search?subtree=&subtree_id=&country%5B%5D=603&geo-resident=true&geo-likely-occurs=true&geo-endemic=true&geo-introduced=true&search_type=count'

page_content=read_html(base_url)


# Find all species blocks
species_blocks <- page_content %>%
  html_nodes("div.Species")  # Adjust this selector if necessary

# Initialize a list to store the data for each species
species_data <- list()

# Loop through each species block and extract both species name and location data
for (i in seq_along(species_blocks)) {
  # Extract the species name (assumed to be inside an <a> tag within the species block)
  species_name <- species_blocks[i] %>%
    html_node("a") %>%
    html_text() %>%
    trimws()
  
  # Extract location data (assumed to be inside <div> with class "highlight")
  location_nodes <- species_blocks[i] %>%
    html_nodes("div.highlight") %>%
    html_text() %>%
    trimws()

  # Initialize variables to store locations by type
  natural_resident <- character(0)
  introduced <- character(0)
  endemic <- character(0)
  likely <- character(0)
  
  # Loop through each location node to check if it matches the types
  for (location in location_nodes) {
    location_cleaned <- str_replace_all(location, "\n", " ") %>%
      str_replace_all("\\s+", " ") %>%
      str_trim()

    # Check for keywords and assign locations accordingly
    if (grepl("Natural Resident", location_cleaned)) {
      natural_resident <- c(natural_resident, str_extract(location_cleaned, "(?<=Natural Resident: ).*"))
    }
    
    if (grepl("Introduced", location_cleaned)) {
      introduced <- c(introduced, str_extract(location_cleaned, "(?<=Introduced: ).*"))
    }
    
    if (grepl("Endemic", location_cleaned)) {
      endemic <- c(endemic, str_extract(location_cleaned, "(?<=Endemic: ).*"))
    }
    
    if (grepl("Likely", location_cleaned)) {
      likely <- c(likely, str_extract(location_cleaned, "(?<=Likely: ).*"))
    }
  }

  # Combine the extracted data into a data frame for each species
  species_data[[i]] <- data.frame(
    Species = species_name,
    Natural_Resident = paste(natural_resident, collapse = "; "),  # Join all locations if there are multiple
    Introduced = paste(introduced, collapse = "; "),
    Endemic = paste(endemic, collapse = "; "),
    Likely = paste(likely, collapse = "; "),
    stringsAsFactors = FALSE
  )
}

# Combine all species data into a single data frame
species_df <- bind_rows(species_data)


# Clean up any residual issues with whitespace
species_df <- species_df %>%
  mutate(
    Species = str_trim(Species),
    Natural_Resident = str_trim(Natural_Resident),
    Introduced = str_trim(Introduced),
    Endemic = str_trim(Endemic),
    Likely = str_trim(Likely)
  )

# View the final data frame
print(species_df)

# View the cleaned data frame
View(species_df)

USAAmphib<-species_df
write.csv(USAAmphib,'USAAmphibOnly.csv')
```



###Global Dataframe

URL is too big to include all data together so split into A through M country names and N through Z country names.

IT WORKS! DO NOT TOUCH THIS BLOCK
```{r loop data pull}

# URL for A through M Countries

base_url='https://amphibiansoftheworld.amnh.org/content/search?subtree=&subtree_id=&country%5B%5D=614&country%5B%5D=389&country%5B%5D=445&country%5B%5D=388&country%5B%5D=302&country%5B%5D=153&country%5B%5D=182&country%5B%5D=414&country%5B%5D=305&country%5B%5D=541&country%5B%5D=555&country%5B%5D=584&country%5B%5D=219&country%5B%5D=527&country%5B%5D=374&country%5B%5D=191&country%5B%5D=457&country%5B%5D=232&country%5B%5D=477&country%5B%5D=440&country%5B%5D=273&country%5B%5D=569&country%5B%5D=593&country%5B%5D=333&country%5B%5D=294&country%5B%5D=607&country%5B%5D=442&country%5B%5D=194&country%5B%5D=347&country%5B%5D=450&country%5B%5D=478&country%5B%5D=373&country%5B%5D=269&country%5B%5D=471&country%5B%5D=563&country%5B%5D=588&country%5B%5D=188&country%5B%5D=521&country%5B%5D=583&country%5B%5D=253&country%5B%5D=505&country%5B%5D=599&country%5B%5D=266&country%5B%5D=160&country%5B%5D=606&country%5B%5D=260&country%5B%5D=436&country%5B%5D=330&country%5B%5D=303&country%5B%5D=564&country%5B%5D=520&country%5B%5D=386&country%5B%5D=321&country%5B%5D=204&country%5B%5D=419&country%5B%5D=367&country%5B%5D=200&country%5B%5D=554&country%5B%5D=534&country%5B%5D=162&country%5B%5D=301&country%5B%5D=508&country%5B%5D=437&country%5B%5D=163&country%5B%5D=585&country%5B%5D=581&country%5B%5D=262&country%5B%5D=482&country%5B%5D=551&country%5B%5D=601&country%5B%5D=611&country%5B%5D=472&country%5B%5D=316&country%5B%5D=1132348&country%5B%5D=243&country%5B%5D=245&country%5B%5D=514&country%5B%5D=490&country%5B%5D=509&country%5B%5D=1132349&country%5B%5D=155&country%5B%5D=307&country%5B%5D=161&country%5B%5D=486&country%5B%5D=310&country%5B%5D=1132350&country%5B%5D=278&country%5B%5D=387&country%5B%5D=441&country%5B%5D=166&country%5B%5D=1264363&country%5B%5D=349&country%5B%5D=223&country%5B%5D=513&country%5B%5D=1132351&country%5B%5D=206&country%5B%5D=365&country%5B%5D=544&country%5B%5D=203&country%5B%5D=196&country%5B%5D=1132352&country%5B%5D=590&country%5B%5D=362&country%5B%5D=552&country%5B%5D=494&country%5B%5D=537&country%5B%5D=846210&country%5B%5D=1132355&country%5B%5D=573&country%5B%5D=435&country%5B%5D=451&country%5B%5D=181&country%5B%5D=454&country%5B%5D=589&country%5B%5D=600&country%5B%5D=491&country%5B%5D=498&country%5B%5D=561&country%5B%5D=344&country%5B%5D=1132353&country%5B%5D=208&country%5B%5D=187&country%5B%5D=267&country%5B%5D=1080969&country%5B%5D=1080968&country%5B%5D=190&country%5B%5D=202&country%5B%5D=210&country%5B%5D=352&country%5B%5D=574&country%5B%5D=205788&country%5B%5D=251&country%5B%5D=525&country%5B%5D=549&country%5B%5D=480&country%5B%5D=277&country%5B%5D=343&country%5B%5D=341&country%5B%5D=424&country%5B%5D=400&country%5B%5D=375&geo-resident=true&geo-likely-occurs=true&geo-endemic=true&geo-introduced=true&search_type=count'

page_content=read_html(base_url)


# Find all species blocks
species_blocks <- page_content %>%
  html_nodes("div.Species")  # Adjust this selector if necessary

# Initialize a list to store the data for each species
species_data <- list()

# Loop through each species block and extract both species name and location data
for (i in seq_along(species_blocks)) {
  # Extract the species name (assumed to be inside an <a> tag within the species block)
  species_name <- species_blocks[i] %>%
    html_node("a") %>%
    html_text() %>%
    trimws()
  
  # Extract location data (assumed to be inside <div> with class "highlight")
  location_nodes <- species_blocks[i] %>%
    html_nodes("div.highlight") %>%
    html_text() %>%
    trimws()

  # Initialize variables to store locations by type
  natural_resident <- character(0)
  introduced <- character(0)
  endemic <- character(0)
  likely <- character(0)
  
  # Loop through each location node to check if it matches the types
  for (location in location_nodes) {
    location_cleaned <- str_replace_all(location, "\n", " ") %>%
      str_replace_all("\\s+", " ") %>%
      str_trim()

    # Check for keywords and assign locations accordingly
    if (grepl("Natural Resident", location_cleaned)) {
      natural_resident <- c(natural_resident, str_extract(location_cleaned, "(?<=Natural Resident: ).*"))
    }
    
    if (grepl("Introduced", location_cleaned)) {
      introduced <- c(introduced, str_extract(location_cleaned, "(?<=Introduced: ).*"))
    }
    
    if (grepl("Endemic", location_cleaned)) {
      endemic <- c(endemic, str_extract(location_cleaned, "(?<=Endemic: ).*"))
    }
    
    if (grepl("Likely", location_cleaned)) {
      likely <- c(likely, str_extract(location_cleaned, "(?<=Likely: ).*"))
    }
  }

  # Combine the extracted data into a data frame for each species
  species_data[[i]] <- data.frame(
    Species = species_name,
    Natural_Resident = paste(natural_resident, collapse = "; "),  # Join all locations if there are multiple
    Introduced = paste(introduced, collapse = "; "),
    Endemic = paste(endemic, collapse = "; "),
    Likely = paste(likely, collapse = "; "),
    stringsAsFactors = FALSE
  )
}

# Combine all species data into a single data frame
species_df <- bind_rows(species_data)


# Clean up any residual issues with whitespace
species_df <- species_df %>%
  mutate(
    Species = str_trim(Species),
    Natural_Resident = str_trim(Natural_Resident),
    Introduced = str_trim(Introduced),
    Endemic = str_trim(Endemic),
    Likely = str_trim(Likely)
  )

# View the final data frame
print(species_df)

# View the cleaned data frame
View(species_df)

AtoM_Amphib<-species_df

```

```{r N-Z data}
# URL for N through Z Countries

base_url='https://amphibiansoftheworld.amnh.org/content/search?subtree=&subtree_id=&country%5B%5D=444&country%5B%5D=392&country%5B%5D=193&country%5B%5D=613&country%5B%5D=370&country%5B%5D=312&country%5B%5D=235&country%5B%5D=174&country%5B%5D=556&country%5B%5D=562&country%5B%5D=511&country%5B%5D=283&country%5B%5D=582&country%5B%5D=1132354&country%5B%5D=492&country%5B%5D=568&country%5B%5D=432&country%5B%5D=167&country%5B%5D=591&country%5B%5D=540&country%5B%5D=175&country%5B%5D=497&country%5B%5D=660345&country%5B%5D=378&country%5B%5D=241&country%5B%5D=578&country%5B%5D=272&country%5B%5D=288&country%5B%5D=669534&country%5B%5D=533&country%5B%5D=239&country%5B%5D=280&country%5B%5D=518&country%5B%5D=669571&country%5B%5D=270&country%5B%5D=248&country%5B%5D=597&country%5B%5D=306&country%5B%5D=617&country%5B%5D=185&country%5B%5D=577&country%5B%5D=506&country%5B%5D=461&country%5B%5D=1173931&country%5B%5D=323&country%5B%5D=438&country%5B%5D=244123&country%5B%5D=393&country%5B%5D=337&country%5B%5D=231&country%5B%5D=363&country%5B%5D=332&country%5B%5D=502&country%5B%5D=615&country%5B%5D=173&country%5B%5D=150&country%5B%5D=335&country%5B%5D=495&country%5B%5D=448&country%5B%5D=325&country%5B%5D=358&country%5B%5D=456&country%5B%5D=545&country%5B%5D=281&country%5B%5D=425&country%5B%5D=183&country%5B%5D=309&country%5B%5D=324&country%5B%5D=592&country%5B%5D=238&country%5B%5D=460&country%5B%5D=559&country%5B%5D=603&country%5B%5D=429&country%5B%5D=523&country%5B%5D=224&country%5B%5D=328&country%5B%5D=229&country%5B%5D=158&country%5B%5D=397&country%5B%5D=52023&country%5B%5D=399&country%5B%5D=247&country%5B%5D=570&country%5B%5D=211&country%5B%5D=532&country%5B%5D=242&country%5B%5D=618&country%5B%5D=380&country%5B%5D=314&country%5B%5D=255&country%5B%5D=595&country%5B%5D=351&country%5B%5D=449&country%5B%5D=430&country%5B%5D=327&country%5B%5D=531&country%5B%5D=176&country%5B%5D=296&country%5B%5D=406&country%5B%5D=215&country%5B%5D=402&country%5B%5D=353&country%5B%5D=408&country%5B%5D=168&country%5B%5D=195&country%5B%5D=519&country%5B%5D=149&country%5B%5D=237&country%5B%5D=510&country%5B%5D=395&country%5B%5D=483&country%5B%5D=360&country%5B%5D=489&country%5B%5D=455&country%5B%5D=439&country%5B%5D=258&country%5B%5D=547&country%5B%5D=339&country%5B%5D=304&country%5B%5D=216&country%5B%5D=284&country%5B%5D=264&country%5B%5D=538&country%5B%5D=274&country%5B%5D=484&country%5B%5D=515&country%5B%5D=609&country%5B%5D=236&country%5B%5D=384&country%5B%5D=322&country%5B%5D=468655&country%5B%5D=300&country%5B%5D=371&country%5B%5D=499&country%5B%5D=366&geo-resident=true&geo-likely-occurs=true&geo-endemic=true&geo-introduced=true&search_type=count'

page_content=read_html(base_url)


# Find all species blocks
species_blocks <- page_content %>%
  html_nodes("div.Species")  # Adjust this selector if necessary

# Initialize a list to store the data for each species
species_data <- list()

# Loop through each species block and extract both species name and location data
for (i in seq_along(species_blocks)) {
  # Extract the species name (assumed to be inside an <a> tag within the species block)
  species_name <- species_blocks[i] %>%
    html_node("a") %>%
    html_text() %>%
    trimws()
  
  # Extract location data (assumed to be inside <div> with class "highlight")
  location_nodes <- species_blocks[i] %>%
    html_nodes("div.highlight") %>%
    html_text() %>%
    trimws()

  # Initialize variables to store locations by type
  natural_resident <- character(0)
  introduced <- character(0)
  endemic <- character(0)
  likely <- character(0)
  
  # Loop through each location node to check if it matches the types
  for (location in location_nodes) {
    location_cleaned <- str_replace_all(location, "\n", " ") %>%
      str_replace_all("\\s+", " ") %>%
      str_trim()

    # Check for keywords and assign locations accordingly
    if (grepl("Natural Resident", location_cleaned)) {
      natural_resident <- c(natural_resident, str_extract(location_cleaned, "(?<=Natural Resident: ).*"))
    }
    
    if (grepl("Introduced", location_cleaned)) {
      introduced <- c(introduced, str_extract(location_cleaned, "(?<=Introduced: ).*"))
    }
    
    if (grepl("Endemic", location_cleaned)) {
      endemic <- c(endemic, str_extract(location_cleaned, "(?<=Endemic: ).*"))
    }
    
    if (grepl("Likely", location_cleaned)) {
      likely <- c(likely, str_extract(location_cleaned, "(?<=Likely: ).*"))
    }
  }

  # Combine the extracted data into a data frame for each species
  species_data[[i]] <- data.frame(
    Species = species_name,
    Natural_Resident = paste(natural_resident, collapse = "; "),  # Join all locations if there are multiple
    Introduced = paste(introduced, collapse = "; "),
    Endemic = paste(endemic, collapse = "; "),
    Likely = paste(likely, collapse = "; "),
    stringsAsFactors = FALSE
  )
}

# Combine all species data into a single data frame
species_df <- bind_rows(species_data)


# Clean up any residual issues with whitespace
species_df <- species_df %>%
  mutate(
    Species = str_trim(Species),
    Natural_Resident = str_trim(Natural_Resident),
    Introduced = str_trim(Introduced),
    Endemic = str_trim(Endemic),
    Likely = str_trim(Likely)
  )

# View the final data frame
print(species_df)

# View the cleaned data frame
View(species_df)

NtoZ_Amphib<-species_df

```
Combine dataframes 

```{r combine}


GlobalAmphib<-rbind(AtoM_Amphib,NtoZ_Amphib)
write.csv(GlobalAmphib,'GlobalAmphib.csv')


```

#Mapping
Using IUCN shapefiles: https://damariszurell.github.io/EEC-MGC/a4_RangeMaps.html

```{r Maps Download}
library(terra)
library(ggplot2)
library(maps)
library(ggspatial)
library(sf)

# Load the shape file
caudata<- terra::vect("/Users/jadyns/Documents/JadynS/UNC/Courses/Hybridization/IUCN Range Maps/CAUDATA/CAUDATA.shp")

anura1<-terra::vect("/Users/jadyns/Documents/JadynS/UNC/Courses/Hybridization/IUCN Range Maps/ANURA/ANURA_PART1.shp")

anura2<-terra::vect("/Users/jadyns/Documents/JadynS/UNC/Courses/Hybridization/IUCN Range Maps/ANURA/ANURA_PART2.shp")

gymnophiona<-terra::vect("/Users/jadyns/Documents/JadynS/UNC/Courses/Hybridization/IUCN Range Maps/GYMNOPHIONA/GYMNOPHIONA.shp")

shapefile_list<-list(caudata,anura1,anura2,gymnophiona)

all_amphib <- do.call(rbind, shapefile_list)

#TO map with ggplot use the simple features package. This takes a while. Be patient :)
all_amphib_sf<-st_as_sf(all_amphib)
```

```{r Spatial Manipulation}
#Check spatial specifics
class(all_amphib_sf)
names(all_amphib_sf)
all_amphib_sf$sci_name<-as.factor(all_amphib_sf$sci_name)

#generate a world map
world_map <- map_data("world")

#Filter by genus or species of interest. This is an example

species_ranges_genus <- all_amphib_sf %>% 
  filter(genus %in% c("Plethodon"))

#Manual Example

unique(all_amphib_sf$legend)

Alytesdickhilleni<-all_amphib_sf %>% 
  filter(sci_name %in% c("Alytes dickhilleni"))

Alytesobstetricans<-all_amphib_sf %>% 
  filter(sci_name %in% c("Alytes obstetricans"))

#Map relevant ranges 
ggplot() +
  geom_map(data = world_map, map = world_map, aes(x = long, y = lat, map_id = region), 
           color = "black", fill = "lightgrey", size = 0.2) +
  geom_sf(data = Alytesdickhilleni, aes(fill = sci_name), color = "transparent",alpha=0.5) +
  theme_minimal() +
  ggtitle("Species Range Map") +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_text(size = 10),  # Adjust legend title size
    legend.text = element_text(size = 8),    # Adjust legend text size
    legend.key.size = unit(0.5, "cm"),      # Adjust legend key size
    legend.position = "right"                # Move the legend to the right of the plot
  ) +
  coord_sf(datum = NA)  # Removes the default coordinate system grid

```

#Range size

###By Species
```{r Range Size}

#Upload overlap data

# Read the data
AmphibAreaList <- read.csv("/Users/jadyns/Documents/JadynS/UNC/Courses/Hybridization/GlobalAmphibHybridArea.csv")

# Initialize vectors to store the results
ResidentArea <- numeric(nrow(AmphibAreaList))
IntroducedArea <- numeric(nrow(AmphibAreaList))
OverlapArea <- numeric(nrow(AmphibAreaList))

# Create a flag column (default is 'No Issue')
AmphibAreaList$IssueFlag <- "No Issue"

# Get unique species
IntroducedVect <- unique(AmphibAreaList$Introduced)
ResidentVect <- unique(AmphibAreaList$Resident)

combined_list <- unique(c(IntroducedVect, ResidentVect))

# Check for species name matches between your list and all_amphib_sf
species_name_issue <- combined_list[!combined_list %in% all_amphib_sf$sci_name]

# Flag rows where the species names in AmphibAreaList don't match
AmphibAreaList$IssueFlag[AmphibAreaList$Resident %in% species_name_issue | AmphibAreaList$Introduced %in% species_name_issue] <- "Species Name Mismatch"


# Loop through the rows of the data frame
for (i in 1:nrow(AmphibAreaList)) {
  # Get the species names for the current row
  ResidentSpeciesName <- AmphibAreaList$Resident[i]
  IntroducedSpeciesName <- AmphibAreaList$Introduced[i]
  
  # Filter the 'all_amphib_sf' data frame for the species 1 and species 2
  ResidentRange <- all_amphib_sf %>% 
    filter(sci_name == ResidentSpeciesName)
  
  IntroducedRange <- all_amphib_sf %>% 
    filter(sci_name == IntroducedSpeciesName)
  
  # Ensure both species ranges have valid geometries
  if (nrow(ResidentRange) == 0 | nrow(IntroducedRange) == 0) {
    AmphibAreaList$IssueFlag[i] <- "No Data Available"
    next  # Skip if no data is found for either species
  }

  # Clean and simplify geometries if necessary
  ResidentRange <- st_make_valid(ResidentRange)
  IntroducedRange <- st_make_valid(IntroducedRange)
  
  # Check if any geometries are empty (now using `any(st_is_empty(...))`)
  if (any(st_is_empty(ResidentRange$geometry)) | any(st_is_empty(IntroducedRange$geometry))) {
    AmphibAreaList$IssueFlag[i] <- "Empty Geometry"
    next  # Skip if either geometry is empty
  }

  # Ensure CRS consistency and use an appropriate CRS for area calculations (e.g., UTM)
  if (st_crs(ResidentRange) != st_crs(IntroducedRange)) {
    IntroducedRange <- st_transform(IntroducedRange, st_crs(ResidentRange))
  }

  # Optional: Transform both to a projected CRS for accurate area calculation
  ResidentRange <- st_transform(ResidentRange, crs = 3395)  # Example: WGS 84/Pseudo Mercator uses meters as area unit
  IntroducedRange <- st_transform(IntroducedRange, crs = 3395)

  # Calculate the areas for Resident and Introduced species ranges
  ResidentArea[i] <- st_area(ResidentRange) %>% sum()
  IntroducedArea[i] <- st_area(IntroducedRange) %>% sum()

  # Check for extremely small or large areas
  if (ResidentArea[i] < 0 | IntroducedArea[i] < 0) {
    AmphibAreaList$IssueFlag[i] <- "Negative Area"
    next
  }
  if (ResidentArea[i] == 0 | IntroducedArea[i] == 0) {
    AmphibAreaList$IssueFlag[i] <- "Zero Area"
    next
  }

  # Calculate the overlap area between Resident and Introduced species ranges
  tryCatch({
    OverlapRange <- st_intersection(ResidentRange, IntroducedRange)
    OverlapArea[i] <- st_area(OverlapRange) %>% sum()

    # Check for zero or negative overlap
    if (OverlapArea[i] <= 0) {
      AmphibAreaList$IssueFlag[i] <- "Zero or Negative Overlap"
    }
  }, error = function(e) {
    message("Error with row ", i, ": ", e$message)
    AmphibAreaList$IssueFlag[i] <- "Intersection Error"
    next
  })
}

# Add the calculated areas to the original data frame
AmphibAreaList$ResidentArea <- ResidentArea
AmphibAreaList$IntroducedArea <- IntroducedArea
AmphibAreaList$OverlapArea <- OverlapArea
AmphibAreaList$IntroducedPerCentOverlap<-OverlapArea/ResidentArea *100

# Convert the 'value' column to non-scientific format
AmphibAreaList$IntroducedPerCentOverlap <- format(AmphibAreaList$IntroducedPerCentOverlap, scientific = FALSE)
# Print the updated data frame with the new columns
print(AmphibAreaList)

write.csv(AmphibAreaList,"AmphibAreaList.csv")
```

###Overlapping Range

(I'm still working on this one...)
```{R overlapping range}

